A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-09-08 docs/results_JAMwithNFW_v1.01.tex
Finished discussion on stellar mass-to-light ratios.
Started writing discussion of possible minor merger.

\subsection{JAM with a NFW Dark Matter Halo}

@@

\paragraph{Modelling.} The full set of fit parameters is $(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta)$ [TO DO: $\beta$ or $\beta_0$???] [TO DO: Einheitlich $\Upsilon_{I,*}$????], where $\beta$ is the constant velocity anisotropy parameter (see \S\ref{sec:model_JAM}). We will investigate this parameter space with a MCMC \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For an Chabrier IMF stellar population they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate [TO DO: introduce somewhere], we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Eq. (\ref{eq:Maccio08}) as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{equation*}
p(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta) = \frac{1}{\ln\left( 10 R_{s,NFW}\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{equation*}
where the factor $1/\ln\left( 10 R_{s,NFW}\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ [TO DO: introduce concentration] to $(v_{200},R_{s,NFW})$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5$ arcsec, approximately within the effective half-light radius $R_\text{eff} = 2.6$ arcsec  [TO DO: make sure R' everywhere ???]. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit.

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The best fit parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies; the sample with the highest probability is pegged at the lower limit of the prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. [TO DO: Rewrite. Not very good.]

The first panel in Figure \ref{fig:modelB4_vrot} shows the mean best fit values from Table \ref{tab:modelB4_bestfit}.

We also fitted a model with a cored logarithmic dark matter halo. The cored halo models are in general slightly less massive than the NFW halo and therefore fitt the outer regions of J1331 better. However, the density profile of the cored halo as well as the I-band light distribution within the plane drop as $\rho(r) \propto r^{-2}$, there is a strong degeneracy between the stellar mass and the dark matter. Overall we were not able to give tight constraints on the cored logarithmic halo.

@@

\begin{table}
\centering
\begin{tabular}{llrcl}
\hline
stellar I-band mass-to-light ratio & $\Upsilon_{I,*}$ & 4.2 & $\pm$ & 0.2\\
velocity anisotropy & $\beta_z$ & -0.4 & $\pm$ & 0.1 \\
NFW halo scale length & $R_\text{s,NFW}$ [kpc] & 40 & $\pm$ & 20\\
NFW halo virial velocity & $v_{200}$ [km/s] & 240 & $\pm$ & 40\\
NFW halo concentration & $c_{200}$ & 8 & $\pm$ & 2 \\
NFW halo mass & $M_{200}$ [$10^{12} M_\odot$] & 5 & $\pm$ & 2\\
\hline
\end{tabular}
\caption{Summary of the best fit parameters of the JAM model with NFW halo from the MCMC exploration in Figure \ref{fig:modelB4_triangle}. The halo mass and concentration are calculated from the the best fit $R_\text{s,NFW}$ and $v_{200}$.}
\label{tab:modelB4_bestfit}
\end{table}

>>>>

\subsection{JAM with a NFW Dark Matter Halo} \label{sec:results_JAM_NFW}

@@

\paragraph{Modelling.} The full set of fit parameters is $(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta)$ [TO DO: $\beta$ or $\beta_0$???] [TO DO: Einheitlich $\Upsilon_{I,*}$????], where $\beta$ is the constant velocity anisotropy parameter (see \S\ref{sec:model_JAM}). We will investigate this parameter space with a MCMC \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For an Chabrier IMF stellar population they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate [TO DO: introduce somewhere], we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Eq. (\ref{eq:Maccio08}) as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{eqnarray*}
&&p(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta) \\
&&= \frac{1}{\ln\left( 10 R_{s,NFW}\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{eqnarray*}
where the factor $1/\ln\left( 10 R_{s,NFW}\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ [TO DO: introduce concentration] to $(v_{200},R_{s,NFW})$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5$ arcsec, approximately within the effective half-light radius $R_\text{eff} = 2.6$ arcsec  [TO DO: make sure R' everywhere ???]. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit. [TO DO: Didn't include the results from the Moster 2010 relation, because they where quite close to the Dutton 2010 relation anyway. Or should I?]

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $R_\text{s,NFW}$) than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies: The sample with the highest probability is pegged at the lower limit of the unifrom prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. There are slight covariances between $\Upsilon_{I,*}$,  $R_\text{s,NFW}$ and $\beta_z$: The smaller the dark matter contribution in the center (i.e. the larger $\Upsilon_{I,*}$ [TO DO: make sure, mass-to-light ratio is everywhere named properly]), the less concentrated is the halo (i.e. the larger $R_\text{s,NFW}$), the more velocity anisotropy is needed to reproduce the central dip. As the effect of $v_{200}$ is mostly at larger radii, this parameter doesn't show any covariances - but is also mostly constrained by the prior. The $v_\text{rms}$ curve corresponding to the mean best fit values from Table \ref{tab:modelB4_bestfit} is shown in the first panel of Figure \ref{fig:modelB4_vrot}. Figure \ref{fig:modelB4_models} illustrates the range of parameters fitting the data according to the full posterior probability function. The model fits the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and is also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6$ arcsec at all and seems to be therefore overall too massive.\\

We also fitted a model with a cored logarithmic dark matter halo. The cored halo models are in general slightly less massive than the NFW halo and therefore fit the outer regions of J1331 better. However, the density profile of the cored halo as well as the I-band light distribution within the plane drop as $\rho(r) \propto r^{-2}$, there is a strong degeneracy between the stellar mass and the dark matter. Overall we were not able to give tight constraints on the cored logarithmic halo.

@@

\begin{table*}
\centering
\begin{tabular}{llrcl}
\hline
stellar I-band mass-to-light ratio & $\Upsilon_{I,*}$ & 4.2 & $\pm$ & 0.2\\
velocity anisotropy & $\beta_z$ & -0.4 & $\pm$ & 0.1 \\
NFW halo scale length & $R_\text{s,NFW}$ [kpc] & 40 & $\pm$ & 20\\
NFW halo virial velocity & $v_{200}$ [km/s] & 240 & $\pm$ & 40\\
NFW halo concentration & $c_{200}$ & 8 & $\pm$ & 2 \\
NFW halo mass & $M_{200}$ [$10^{12} M_\odot$] & 5 & $\pm$ & 2\\
\hline
\end{tabular}
\caption{Summary of the best fit parameters of the JAM model with NFW halo from the MCMC exploration in Figure \ref{fig:modelB4_triangle}. The halo mass and concentration are calculated from the the best fit $R_\text{s,NFW}$ and $v_{200}$.}
\label{tab:modelB4_bestfit}
\end{table*}
